Black-hole ringdown as a probe of higher-curvature gravity theories

### 2022-02-15 papers/paper_alt_theor_bounds/paper_alt_theor_bounds.tex
hs: started super terse text on data analysis

\documentclass[twocolumn,prd,aps,superscriptaddress,preprintnumbers,tightenlines,showpacs,nofootinbib,eqsecnum,amsfonts,amsmath,longbibliography]{revtex4-1}
\usepackage{epsfig}
\usepackage{graphics}
\usepackage{graphicx}
\usepackage[dvipsnames]{xcolor}
\usepackage{bm}
\usepackage{txfonts}
\usepackage{natbib}
\usepackage{amssymb}
\usepackage{xspace}
\usepackage[normalem]{ulem} % To get strikethrough (\sout)
\usepackage[colorlinks]{hyperref}
\usepackage[caption=false]{subfig}
\usepackage{booktabs}
\usepackage{url}
\usepackage{float}
\usepackage[bottom]{footmisc}
\usepackage{lineno}
\usepackage{mathrsfs}
\usepackage{makecell}
\usepackage{microtype}

@@

% ---------------------------
% CONCRETE EXAMPLE OF MAPPING
% ---------------------------
Let us show one concrete example of how this mapping is done.
%
In Ref.~\cite{Wagle:2021tam} the QNMs of slow-rotating BHs in dCS gravity were calculated
and it was found that the axial gravitational modes have their damping time increased as we
increase the Chern-Simons coupling $\ell_{\rm CS}$ at fixed spin $\chi$ of the BH.
%
Having selected this mode we can now obtain $\delta\omega^{(i)}$ and $\delta\tau^{(i)}$ as follows.
%
Using the fitting formula Eq.~(54a) of~\cite{Wagle:2021tam}, namely,
%
\begin{equation}
    M \omega^{\rm CS}_{220} = c_{1} + c_{2} \kappa \zeta + (c_{3} + c_{4} \kappa \zeta) \, (1 - \chi)^{c_{5} + c_{6} \kappa \zeta},
    \label{eq:omega_dsc_eg}
\end{equation}
%
and similarly for the imaginary part ${\rm Im}(\sigma_{220}) = \omega^{\rm CS}_{220} = 1 / \tau^{\rm CS}_{220}$.
%
Here $\kappa = 1/(16 \pi)$, $\zeta = \ell_{\rm CS}^{4} / (M^4 \kappa)$ [such that $\kappa \zeta = (\ell_{\rm CS} / M)^4$]
and $c_{i}$ are fitting coefficients.
%
In the context of a binary BH merger, $M$ is interpreted as the remnant mass $\Mf$.

We can now expand Eq.~\eqref{eq:omega_dsc_eg} to leading-orders in $\chi$ and $\ell_{\rm CS}$, and
gather the terms proportional to $\ell_{\rm CS}$.
%
Doing so, we find,
%
\begin{align}
    \Mf \, \omega^{\rm CS}_{220} &=
    \left[ 0.3722 + 1.1945 (\ell_{\rm CS} / \Mf)^4 \right]
    \nonumber \\
    &\quad + \left[0.1861 + 5.1828 (\ell_{\rm CS} / \Mf)^4 \right] \, \chi\,,
\end{align}
%
where we made use of the numerical value of the coefficients $c_{i}$ and,
reassuringly, the nonrotating and linear-in-spin coefficients above agree with those
of~\cite{Maselli:2019mjd} to xx\%.
%
We now isolate the dCS corrections and compare against Eq.~\eqref{eq:qnm_nongr_iso}.
%
Doing so, we can make the identifications,
%
\begin{equation}
p_{\rm cs} = 4\,, \quad \delta \omega^{(0)}_{\rm cs} = 3.1964\,, \quad \delta \omega^{(1)}_{\rm cs} = 41.199\,.
\label{eq:cs_omega_coefs}
\end{equation}
%
The same procedure can be made for $\tau^{(i)}$, which results in,
%
\begin{equation}
\delta \tau^{(0)}_{\rm cs} = 6.3619\,, \quad \delta \tau^{(1)}_{\rm cs} = 794.66\,.
\label{eq:cs_tau_coefs}
\end{equation}
%
which completes the set of fixed parameters beyond-GR parameters in the ringdown of the \pSEOB waveform model.
%
As we said before, the only free beyond-GR parameter left is now $\ell_{\rm CS}$.

@@

\hs{Say that we use GW150914~\cite{LIGOScientific:2016aoc} and GW200129~\cite{LIGOScientific:2021djp} and justify why.}

@@

>>>>

\documentclass[twocolumn,prd,aps,superscriptaddress,preprintnumbers,tightenlines,showpacs,nofootinbib,eqsecnum,amsfonts,amsmath,longbibliography]{revtex4-1}
\usepackage{epsfig}
\usepackage{graphics}
\usepackage{graphicx}
\usepackage[dvipsnames]{xcolor}
\usepackage{bm}
\usepackage{txfonts}
% \usepackage{newtxtext}
% \usepackage{newtxmath}
\usepackage{natbib}
% \usepackage{amssymb}
\usepackage{xspace}
\usepackage[normalem]{ulem} % To get strikethrough (\sout)
\usepackage[colorlinks]{hyperref}
\usepackage[caption=false]{subfig}
\usepackage{booktabs}
\usepackage{url}
\usepackage{float}
\usepackage[bottom]{footmisc}
\usepackage{lineno}
\usepackage{mathrsfs}
\usepackage{makecell}
\usepackage{microtype}

@@

% ---------------------------
% CONCRETE EXAMPLE OF MAPPING
% ---------------------------
Let us show one concrete example of how this mapping is done.
%
In Ref.~\cite{Wagle:2021tam} the QNMs of slow-rotating BHs in dCS gravity were calculated
and it was found that the axial gravitational modes have their damping time increased as we
increase the Chern-Simons coupling $\ell_{\rm CS}$ at fixed spin $\chi$ of the BH.
%
Having selected this mode we can now obtain $\delta\omega^{(i)}$ and $\delta\tau^{(i)}$ as follows.
%
Using the fitting formula Eq.~(54a) of~\cite{Wagle:2021tam}, namely,
%
\begin{equation}
    M \omega^{\rm CS}_{220} = c_{1} + c_{2} \kappa \zeta + (c_{3} + c_{4} \kappa \zeta) \, (1 - \chi)^{c_{5} + c_{6} \kappa \zeta},
    \label{eq:omega_dsc_eg}
\end{equation}
%
and similarly for the imaginary part ${\rm Im}(\sigma_{220}) = \omega^{\rm CS}_{220} = 1 / \tau^{\rm CS}_{220}$.
%
Here $\kappa = 1/(16 \pi)$, $\zeta = \ell_{\rm CS}^{4} / (M^4 \kappa)$ [thus $\kappa \zeta = (\ell_{\rm CS} / M)^4$]
and $c_{i}$ are fitting coefficients.
%
In the context of a binary BH merger, $M$ is interpreted as the remnant mass $\Mf$.

We can now expand Eq.~\eqref{eq:omega_dsc_eg} to leading-orders in $\chi$ and $\ell_{\rm CS}$, and
gather the terms proportional to $\ell_{\rm CS}$.
%
Doing so, we find,
%
\begin{align}
    \Mf \, \omega^{\rm CS}_{220} &=
    \left[ 0.3722 + 1.1945 (\ell_{\rm CS} / \Mf)^4 \right]
    \nonumber \\
    &\quad + \left[0.1861 + 5.1828 (\ell_{\rm CS} / \Mf)^4 \right] \, \chi\,,
\end{align}
%
where we made use of the numerical value of the coefficients $c_{i}$ and,
reassuringly, the nonrotating and linear-in-spin coefficients above agree with those
of~\cite{Maselli:2019mjd} to xx\%.
%
We now isolate the dCS corrections and compare against Eq.~\eqref{eq:qnm_nongr_iso}.
%
Doing so, we can make the identifications,
%
\begin{equation}
p_{\rm cs} = 4\,, \quad \delta \omega^{(0)}_{\rm cs} = 3.1964\,, \quad \delta \omega^{(1)}_{\rm cs} = 41.199\,.
\label{eq:cs_omega_coefs}
\end{equation}
%
The same procedure can be carried for $\tau^{(i)}$, which results in,
%
\begin{equation}
\delta \tau^{(0)}_{\rm cs} = 6.3619\,, \quad \delta \tau^{(1)}_{\rm cs} = 794.66\,.
\label{eq:cs_tau_coefs}
\end{equation}
%
which completes the set of fixed parameters beyond-GR parameters in the ringdown of the \pSEOB{} waveform model.
%
As we said before, the only free beyond-GR parameter left is now $\ell_{\rm CS}$.

@@

% \hs{Say that we use GW150914~\cite{LIGOScientific:2016aoc} and GW200129~\cite{LIGOScientific:2021djp} and justify why.}

We will consider the events GW150915~\cite{LIGOScientific:2016aoc} and GW200129~\cite{LIGOScientific:2021djp}.
%
They are suitable for ringdown studies due to the high-masses and high signal-to-noise ratios (SNR).

\begin{equation}
P(\bm{\lambda} \vert d, \mathcal{H}) =
\frac{P(\bm{\lambda} \vert \mathcal{H}) P(d \vert \bm{\lambda},\mathcal{H})}{P(d \vert \mathcal{H})}\,,
\label{eq:bayes}
\end{equation}
%
where $d$ is the data and $\bm{\lambda}$ is a set of vectors that fully
characterize the waveform.
%
$P(\bm{\lambda} \vert \mathcal{H})$ is prior probability distribution,
$P(d \vert \bm{\lambda},\mathcal{H})$ the likelihood, and
$P(d \vert \mathcal{H})$ the evidence.

Our set of parameters is $\bm{\lambda} = \bm{\theta}_{\rm GR} \cup \bm{\theta}_{\rm nGR}$, where
$\bm{\theta}_{\rm nGR} = \{\ell_{i}\}$ is a set of beyond-GR coupling constants.
%
As discussed in Sec.~\ref{sec:review_theories} there is a single coupling
constants in the cases of EdGB and dCS theories, but two in the cubic and
quartic EFTofGR.

Assuming Gaussian and stationary noise, we can write the (log) likelihood function as,
%
\begin{equation}
\ln \mathcal{L}(d \vert \bm{\lambda},\mathcal{H}) \propto
- \tfrac{1}{2}
\langle
d - h(\bm{\lambda}) \vert d - h(\bm{\lambda})
\rangle\,,
\end{equation}
%
with noise-weighted inner product $\langle \cdot | \cdot \rangle$ defined as,
%
\begin{equation}
\langle A | B \rangle =
\int_{f_{\rm low}}^{f_{\rm high}} \, \dd f \,
\frac{\tilde{A}^{\ast}(f) \tilde{B}(f) + \tilde{A}(f) \tilde{B}^{\ast}(f)}{S_{n}(f)}
\end{equation}
%
where $\tilde{A}(f)$ is the Fourier transform of $A(t)$, the asterisk
denotes complex conjugation and $S_{n}(f)$ is the power spectrum density
of the detector.

With the posterior distribution function $P(\bm{\lambda} | d, \mathcal{H})$ determined, we can then
marginalize over all nuisance parameters to obtain the posterior
distribution function the $i$-th beyond-GR coupling constants $\ell_{i}$,
%
\begin{equation}
P(\ell_{i} | d, \mathcal{H}) =
\int \dd \xi_{\rm GR} \,
P(\bm{\lambda} | d, \mathcal{H}) \,.
\label{eq:marg_posterior}
\end{equation}
%

@@

\subsection{Einstein-dilaton-Gauss-Bonnet}
\label{sec:results_edgb}

\subsection{Dynamical Chern-Simons}
\label{sec:results_dcs}

\subsection{Effective-field-theory of general relativity}
\label{sec:results_efts}
